100 Years of Social Work Research: A Data Science Perspective

### 2015-01-11 Analysis.Rmd
Updated analyses. Added full Ebsco data.

# Overview of Search Strategy

Date conducted:   January 8, 2014

@@

```{r echo=FALSE}
pi.df <- pi.df %>% filter(
        record != "Journal of Applied Social Sciences" &
        record != "Early Child Development and Care" &
        record != "The Clinical Supervisor" &
        record != "Children and Youth Services Review" &
        record != "General Hospital Psychiatry" &
        record != "Canadian Journal on Aging" &
        record != "Canadian Journal of Community Mental Health" &
        record != "Behavior Modification" &
        record != "Employee Assistance Quarterly" &
        record != "Journal of Applied Behavioral Science" & 
        record != "The Scientific Review of Mental Health Practice: Objective Investigations of Controversial and Unorthodox Claims in Clinical Psychology, Psychiatry, and Social Work" &
        record != "Systems Research and Behavioral Science" & 
        record != "Journal of Applied Rehabilitation Counseling" & 
        record != "Journal of Educational & Psychological Consultation" &
        record != "Journal of Community & Applied Social Psychology" &                                          
        record != "Evaluation and Program Planning" &                                                                  
        record != "Journal of Nonverbal Behavior" &
        record != "Brain and Cognition" &
        record != "Early Education and Development")
```

# What is the overall number and names of journal titles?

@@

## Number of unique journal titles by year

@@

journal.count <- ggplot(n.journals.year, aes(as.numeric(year), y=n, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("year") + 
    ylab("frequency") + 
    ggtitle("Number of Journals by Year") + 
    scale_x_continuous(breaks=seq(1914, 2014, 5))

@@

```{r journal_lifespan, echo=FALSE}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first), 
              year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff) %>%
       mutate(stop = year.diff, event = ifelse(as.numeric(last) != 2014, 1, 0)) %>%
       select(title, stop, event, as.numeric(first))

survival.journals <- survfit(Surv(longest.running$stop+1, longest.running$event) ~ 1)
median.survival <- data.frame(time = c(12,12), quant = c(.5,0))

@@

ggsurv(survival.journals) + 
    #geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    #annotate("segment", x = 18, xend = 12, y = .12, yend = .15, size = .25, arrow =arrow()) +
    #geom_text(x = 29, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year in print", size = 4) +
    ylim(0,1)

@@

plot.article.count <- ggplot(df, aes(as.factor(years), 
                    y = year.count, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("year") + 
    ylab("count") + 
    ggtitle("Number of Studies by Year") +
    scale_x_discrete(breaks=c(seq(1914, 2014, 10))) +
    scale_y_continuous(breaks = c(seq(0, 2000, 250)))

@@

subject.terms <- stringr::str_split(kp.df$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, 
                        function(x) gsub(" ", "", x))))

@@

"social justice"
"evidence-based practice" | "evidence based practice" | "EBP"
"bullying"
"child welfare"

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}

LO.df <- filter(pi.df, attributes == "LO")

subject.terms <- stringr::str_split(LO.df$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, function(x) gsub(" ", "", x))))
subject.terms.unique <- length(unique(subject.terms))

location <- arrange(most.frequent, desc(Freq))

print(subject.terms.l)

print(location)
write.csv(location, "location.csv")

@@

top.locations <- location %>% 
                 select(subject.terms) %>%
                 rename(country = subject.terms) %>%
                 slice(1:5)

location.year.top <- location.year %>% filter(country %in% top.locations$country)
ggplot(location.year.top, aes(x = year, y = n, colour = country )) + geom_line()

top.locations.no.us <- location %>% 
                select(subject.terms) %>%
                rename(country = subject.terms) %>%
                slice(2:7)

location.year.top.no.us <- location.year %>% filter(country %in% top.locations.no.us$country)

ggplot(location.year.top.no.us, aes(x = year, y = n, colour = country )) + geom_line()

@@

# Methodology

```{r methodology, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}

MD.df <- filter(pi.df, attributes == "MD")

methodology <- stringr::str_split(MD.df$record, pattern = ";")
methodology.terms <- unlist(lapply(methodology, function(x) gsub(" ", "", x)))

methodology.table <- as.data.frame(table(methodology.terms))
methodology.table <- arrange(methodology.table, desc(Freq))

@@

## Methodology over time

```{r echo=FALSE}
year <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

methodology <- pi.df %>%
        filter(attributes == "MD") %>%
        select(articleID = articleID, method = record) %>%
        mutate(method = gsub(" ", "", method))

methodology.year <- methodology %>% 
         left_join(year) %>%
         group_by(method, year) %>%
         summarise(n = n()) %>%
         arrange(method, year, n)

methodology.year.top <- slice(methodology.table, 1:5) %>%
         rename(method = methodology.terms) %>%
         mutate(method = gsub(" ", "", method))

methodology.year.reduced <- methodology.year %>%
        filter(method %in% methodology.year.top$method)

ggplot(methodology.year.reduced, aes(x = year, y = n, colour = method)) + 
        geom_line()
```

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE}
n.authors.article <- pi.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

n_authors <- n.authors.article %>% 
        group_by(id) %>%
        summarise(n = n())

ggplot(n_authors, aes(x = factor(n))) + 
    geom_bar() + 
    stat_bin(binwidth=1) +
    xlab("Number of Listed Authors") + 
    ylab("Number of Articles")

summary(n_authors$n)
```

@@

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4}
df.2 <- tbl_df(pi.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

@@

plot.author.count2 <- ggplot(n_authors, aes(as.numeric(year), y=average.n, group=1)) + 
    geom_line(colour="black") + 
    geom_ribbon(aes(ymin = average.n-std.dev, ymax=average.n+std.dev), alpha=.2)

head(n_authors, 20)
tail(n_authors, 20)
plot.author.count2
```

>>>>

Data collection date:  January 8, 2014

@@

Note the exclusion of social work today.

```{r echo=FALSE}
pi.df <- pi.df %>% filter(
        record != "Journal of Applied Social Sciences" &
        record != "Early Child Development and Care" &
        record != "The Clinical Supervisor" &
        record != "Children and Youth Services Review" &
        record != "General Hospital Psychiatry" &
        record != "Canadian Journal on Aging" &
        record != "Canadian Journal of Community Mental Health" &
        record != "Behavior Modification" &
        record != "Employee Assistance Quarterly" &
        record != "Journal of Applied Behavioral Science" & 
        record != "The Scientific Review of Mental Health Practice: Objective Investigations of Controversial and Unorthodox Claims in Clinical Psychology, Psychiatry, and Social Work" &
        record != "Systems Research and Behavioral Science" & 
        record != "Journal of Applied Rehabilitation Counseling" & 
        record != "Journal of Educational & Psychological Consultation" &
        record != "Journal of Community & Applied Social Psychology" &                                          
        record != "Evaluation and Program Planning" &                                                                  
        record != "Journal of Nonverbal Behavior" &
        record != "Brain and Cognition" &
        record != "Early Education and Development" & 
        record != "American Journal of Preventive Medicine" &
        record != "Social Work Today")
```

This section provides further cleaning of retained journals, namely merging two journals with abbreviated titles:

```{r}

pi.df$record <- ifelse(pi.df$record == "Practice", "Practice: Social Work in Action", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 74, 42, pi.df$articleID)

pi.df$record <- ifelse(pi.df$record == "Eugenics & Social Welfare Bull.", "Eugenics and Social Welfare Bulletin", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 28, 32, pi.df$articleID)

@@

@@

journal.count <- ggplot(n.journals.year, aes(as.numeric(year), y=n, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Publication year") + 
    ylab("Number") + 
    ggtitle("Number of Journals Over Time") + 
    scale_x_continuous(breaks=seq(1914, 2014, 5))

@@

```{r journal_lifespan, echo=FALSE}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first), 
              year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff) %>%
       mutate(stop = year.diff, event = ifelse(as.numeric(last) != 2013, 1, 0)) %>%
       select(title, stop, event, as.numeric(first))

survival.journals <- survfit(Surv(longest.running$stop+1, longest.running$event) ~ 1)
median.survival <- data.frame(time = c(24,24), quant = c(.5,0))

@@

ggsurv(survival.journals) + 
    geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    annotate("segment", x = 32, xend = 27, y = .12, yend = .12, size = .25, arrow =arrow()) +
    geom_text(x = 45, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year in print", size = 4) +
    ylim(0,1) +
    xlim(0,100) + 
    xlab("Survival Time (in years)") +
    ylab("Survival Probability") +
    ggtitle("Survival Probability of Social Work Journals")

@@

plot.article.count <- ggplot(df, aes(as.factor(years), 
                    y = year.count, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("year") + 
    ylab("count") + 
    ggtitle("Number of Studies by Year") +
    scale_x_discrete(breaks=c(seq(1914, 2015, 10))) +
    scale_y_continuous(breaks = c(seq(0, 2000, 250)))

@@

subject.terms <- stringr::str_split(kp.df$record, pattern = ";")
subject.terms <- tolower(subject.terms)
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, 
                        function(x) gsub(" ", "", x))))

@@

+ social justice
+ evidence-based practice" | "evidence based practice" | "EBP"
+ bullying
+ child welfare

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}
kp.yr <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

kp <- pi.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keyword = record) %>%
        mutate(keyword = gsub(" ", "", keyword))

kp.year <- kp %>% 
         left_join(kp.yr) %>%
         group_by(keyword, year) %>%
         summarise(n = n()) %>%
         arrange(keyword, year, n)

select.keywords <- c("power", "oppression", "discrimination", "socialjustice", "qualitativeresearch")
select.keywords.df <- kp.year %>% filter(keyword %in% select.keywords)

library(plyr) # Fix code so ply loads before dplyr
key.plot.df <- ddply(select.keywords.df, .(keyword), transform, cumulative = cumsum(n))
detach(package:plyr)

#Fix graph to sort the keywords / colour
ggplot(key.plot.df, aes(x = year, y = cumulative, colour = keyword, group = keyword)) + 
    geom_line() +
    ggtitle("Cumulative Sum of Articles by Keywords") +
    ylab("Cumulative Sum") + 
    xlab("Year") + 
    theme(legend.title=element_blank())

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
LO.df <- filter(pi.df, attributes == "LO") %>%
         group_by(record) %>%
         summarise(Country = unique(record),
                   N = n()) %>%
        arrange(desc(N)) %>% 
        select(-record)

locations <- as.data.frame(table(LO.df))
locations <- arrange(locations, desc(Freq))
locations.unique <- length(unique(locations))

#write.csv(location, "location.csv")

@@

top.locations <- LO.df %>% 
                 select(Country) %>%
                 slice(1:5)

location.year.top <- location.year %>% filter(country %in% top.locations$Country)

ggplot(location.year.top, aes(x = year, y = n, colour = country )) + geom_line() +
    ggtitle("Number of articles by Top 5 Countries")

top.locations.no.us <- LO.df %>% 
                select(Country) %>%
                slice(2:6)

location.year.top.no.us <- location.year %>% filter(country %in% top.locations.no.us$Country)

ggplot(location.year.top.no.us, aes(x = year, y = n, colour = country )) + geom_line() +
    ggtitle("Number of article by Top 5 Countries (excluding United States)")
```

@@

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
n.authors.article <- pi.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

n_authors.1 <- n.authors.article %>% 
        group_by(id) %>%
        summarise(n = n())

n_authors.1t <- n_authors.1 %>% mutate(n = ifelse(n >=13, 13, n))

ggplot(n_authors.1t, aes(x=as.factor(n))) + 
    geom_bar(fill="white", colour="black") + 
    #stat_bin(binwidth=.5) +
    ggtitle("Number of Authors Listed Per Article") + 
    xlab("Author") +
    ylab("Number of Articles") +
    scale_y_continuous(limits = c(0, 15000), breaks = seq(0, 14000, 2000)) + 
    scale_x_discrete(limits = c(as.character(1:12), "13+"))

@@

@@

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
df.2 <- tbl_df(pi.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

@@

plot.author.count2 <- ggplot(n_authors, aes(as.numeric(year), y=average.n, group=1)) +  
    geom_ribbon(aes(ymin = average.n-std.dev, ymax=average.n+std.dev), colour="lightblue", fill="lightblue", alpha=.40) +
    geom_line(colour="black") + 
    geom_segment(aes(x= 1940, y = 5.5, xend = 1935, yend = 5.5), 
                 arrow=arrow(length=unit(0.25, "cm"))) +
    annotate("text", x = 1951, y = 5.5, label = "1 outlier case of \n24 listed authors", size=3.5) +
    ylab("Average number of authors (black) \nStandard deviation (light blue") +
    xlab("Publication year") +
    ggtitle("Average Number of Authors per Article Over Time")

plot.author.count2

#n_authors_unique <- n_authors %>% 
#        left_join(year) %>%
#        group_by(author) %>%
#        summarise(Distinct = n())
